### Writing your own functions

Writing your code into functions is key for writing neat R code and prevents repitition of code. In my experience, young biologists are reluctant to use functions in their code and often when functions are being used they are abused and created to do too much at once. RStudio has several built in functions, and with more packages installed, more functions are available. Throughout this course we have continuously been making use of functions. The **tidyverse** package in itself contain several functions, such as the `mean()`, `sqrt()` and **ggplot2** functions.

A typical example of a function would be `sqrt()`. Where the argument must be a number, and the return value is the square root of that number. Executing a function is called *calling* the function. Here we *call* a built in R function.

```{r, eval=FALSE, purl=FALSE}
b <- sqrt(a)
```

Here, the value of `a` is given to the `sqrt()` function, the `sqrt()` function calculates the square root, and returns the value which is then assigned to the object `b`. This function is very simple, because it takes just one argument.

Now we learn to write our own basic function. The process in which we write any function, regardless of the size, is similar and involves three key steps:

- Make use of the term `function`
- Specify the arguements in that function
- Specify the body of the function

```{r, eval=FALSE}
my_func <- function(arg1, arg2) {
  body
}
```

`my_func` is the function name. This can be any variable name, but as previously mentioned throughout this workshop, we should avoid using names that are used in built in R functions such as `mean`, `length`, *etc*., as this may often lead to confusion and ultimately result in unnessasary errors.

`arg1` and `arg2` are the *arguements* of the function. You can write a function with any number of arguments, and some arguments may have default values specified. 

Above, 'body' denotes the function's body. This is all the code found within the { ... } brackets. The size of this code may vary depending on the required outcome. 

Once we start putting things in functions so that we can re-use them, we need to start testing that those functions are working correctly. To see how to do this, lets write a function. As mentioned earlier some arguments may have default values specified and so here I specify `y` to have a default value of 1. So, in the function call, when no value is assigned to `y` the default value will be 1.

```{r}
add <- function(x , y = 1) {
  x + y
}
```

Here we created a function called `add()`. This function takes two arguments, `x` and `y`, and returns the sum of `x` and `y`. Since `y` already has a default value assigned to it, it is necessary to only specify `x` in the function call; specifying `y` is optional. Since the code was 'activated' by running it through the console, the function is now available, like any other built-in functions in R, for you to use.

Now lets put our function to test:

```{r}
add(15) # When no y value is given it is taken as 1, as specified in the function

add(15, 5)
```

For some practice lets write a function where we calculate the sum of the squares:

```{r}
sum_squares <- function(x, y) {
  x^2 + y^2
}
```

Testing the function

```{r}
sum_squares(9, 20)
```

## Basic plotting using functions

```{r}
library(tidyverse)
chicks <- datasets::ChickWeight

Diet1 <- chicks %>% 
  filter(Diet == "1")
Diet2 <- chicks %>% 
  filter(Diet == "2")
Diet3 <- chicks %>% 
  filter(Diet == "3")
```

```{r}
func_plot <- function(df) {
  plot1 <- df %>% 
    ggplot(aes(x = Time, y = weight)) +
    geom_point() +
    geom_smooth(method = "lm") +
    labs(x = "Days", y = "Weight (gm)", colour = "diet type") 
}
```

```{r}
(diet1 <- func_plot(df = Diet1))
(diet2 <- func_plot(df = Diet2))
(diet3 <- func_plot(df = Diet3))
```

### Summary statistics func.

```{r}
sum_stats <- function(df) {
  new_data <- df %>% 
    summarise(mean_wt = mean(weight),
              sd_wt = sd(weight),
              var_wt = var(weight)) 
}
```

```{r}
(diet1 <- sum_stats(df = Diet1))
(diet2 <- sum_stats(df = Diet2))
(diet3 <- sum_stats(df = Diet3))
```

## Exercise:










