---
title: "cold_spells"
author: "Amieroh Abrahams"
date: "27 January 2019"
output: html_document
---

# Libraries

First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(lubridate)
library(FNN)
library(data.table)
library(heatwaveR)
```

### Load site list SACTN data

Now to get to the data. The first step involves the loading of the site list. The statistical properties of the seawater temperature representing the South African coastline, such as the mean, minimum and maximum temperatures. These values vary among coastal sections due to the influence of the cold Benguala and warm Agulhas currents. Here we will only focus on the temperature data found along the west coast (wc) (i.e. sites influenced by the Benguela current, EBUS). The SACTN dataset comprise of 129 *in situ* coastal seawater temperatures derived from daily measurements over up to 40 years. The SACTN temperature dataset was compiled by measuring coastal temperatures at 129 sites along the coast of South Africa, daily from 1972 until 2017. 

```{r load_files1, include=TRUE}
load("Data/site_list_v4.2.RData")
load("Data/SACTN_daily_v4.2.RData")
```

Now we select only the sites occuring along the west coast. Thereafter, I include all the sites which had a time series of greater than 10 years. 

```{r}
site_list_sub <- site_list %>% 
  filter(coast == "wc")

SACTN_daily_wc <- site_list_sub %>% 
  filter(length > 3650)

upwelling_sites <- SACTN_daily_wc[c(-6, -13),] 

SACTN_temps <- SACTN_daily_v4.2 %>% 
  left_join(site_list[,c(4,13)], by = "index") %>% 
  filter(index %in% upwelling_sites$index)
# save(SACTN_daily_temps, file = "SACTN_daily_temps.RData")

SACTN_split <- SACTN_temps %>% 
  separate(index, into = c("site", "src"), sep = "/", remove = FALSE)

SACTN_split_PN <- SACTN_split %>%
  filter(site == "Port Nolloth") %>% 
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()
```

Detecting MCS within the *in situ* and remotely SST datasets

```{r}
ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
             pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology # Show a portion of the climatology
MCS_events <- res$event[1:5,1:10] # Show some of the cold-spells
# MCS <- MCS[14408:15503,] # Get the years from 2013 - 2015
```

PLotting

```{r}
MCS1974 <- event_line(res, spread = 100, metric = "intensity_cumulative",
           start_date = "1974-01-01", end_date = "1977-12-31")

#ggsave(plot = MCS1974, filename = "MCS1974.png")
```

```{r}
ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
             pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology # Show a portion of the climatology
res$event[1:5,1:10] # Show some of the cold-spells
# MCS <- MCS[14408:15503,] # Get the years from 2013 - 2015

library(ggplot2)

# ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"),
#              pctile = 10)
# res <- detect_event(ts)
# event_line(res, spread = 100, metric = "intensity_cumulative",
#            start_date = "2013-12-31", end_date = "2014-12-31")
# 
# event_line(res, spread = 100, start_date = "2013-12-31",
#            end_date = "2014-12-31", category = TRUE)

MCS_SACTN_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS_SACTN")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("SACTN")
# 
# ggsave(plot = MCS_SACTN_PN, filename = "Figures/event_detect/PN/MCS_SACTN_PN.pNG")
# ts <- ts2clm(SACTN_split_PN, climatologyPeriod = c("1974-01-01", "2016-12-31"), pctile = 10)
# res <- detect_event(ts, coldSpells = TRUE)
# MCS <- res$event
# 
# loli_SACTN_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
#   geom_lolli(aes(colour = intensity_cumulative)) +
#   scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
#   xlab("Date") + ylab("Event duration [days]")
# ggsave(plot = loli_SACTN_PN, filename = "Figures/event_detect/PN/loli_SACTN_PN.pNG")

# ggplot(MCS, aes(x = event_no, y = intensity_max)) +
#   geom_lolli(shape = 5, aes(colour = rate_onset), linetype = "dotted") +
#   scale_color_distiller(palette = "RdYlGn", name = "Rate \nonset") +
#   xlab("Event number") + ylab("Max intensity [degree C]")


```

# Repeat this on all of the datasets but only for PN
# MUR
```{r}
load("~/Documents/Masters_2019/MastersProject/Data/MUR_SST.RData")
MUR_SST <- MUR_SST %>%
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(MUR_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
# MCS <- MCS[14408:15503,]

library(ggplot2)

MCS_MUR_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]")))
# ggsave(plot = MCS_MUR_PN, filename = "Figures/MCS.pNG")

# Here i get the years from 2013 to 2016
MCS <- MCS[3868:4961,]
MCS_MUR_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("MUR")
# ggsave(plot = MCS_MUR_PN, filename = "Figures/event_detect/PN/MCS_MUR_PN.pNG")



ts <- ts2clm(MUR_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$event

loli_MUR_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
  geom_lolli(aes(colour = intensity_cumulative)) +
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
  xlab("Date") + ylab("Event duration [days]")
# ggsave(plot = loli_MUR_PN, filename = "Figures/event_detect/PN/loli_MUR_PN.pNG")
```

#G1SST

```{r}
load("Data/G1SSTsub_SST.RData")

G1SSTsub_SST <- G1SSTsub_SST %>% 
  filter(site == "Port Nolloth") %>% 
  dplyr::rename(temp = nearest1) %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(G1SSTsub_SST, climatologyPeriod = c("1991-09-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
# MCS <- MCS[14408:15503,]

library(ggplot2)

MCS_G1SST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]")))
# ggsave(plot = MCS_MUR_PN, filename = "Figures/MCS.pNG")

# Here i get the years from 2013 to 2016
MCS <- MCS[7797:8888,]
MCS_G1SST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("G1SST")
ggsave(plot = MCS_G1SST_PN, filename = "Figures/event_detect/PN/MCS_G1SST_PN.pNG")



ts <- ts2clm(G1SSTsub_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$event

loli_G1SST_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
  geom_lolli(aes(colour = intensity_cumulative)) +
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
  xlab("Date") + ylab("Event duration [days]")

# ggsave(plot = loli_G1SST_PN, filename = "Figures/event_detect/PN/loli_G1SST_PN.pNG")
```

OISST

```{r}
load("~/Documents/Masters_2019/MastersProject/Data/OISST_sites.RData")
OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  group_by(date) %>% 
  drop_na() %>% 
  filter(site == "Port Nolloth") %>% 
  select(date, temp) %>% 
  dplyr::rename(t = date)

ts <- ts2clm(OISST_sites, climatologyPeriod = c("1981-09-01", "2018-06-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$climatology
# MCS <- MCS[14408:15503,]

library(ggplot2)

MCS_OISST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]")))
# ggsave(plot = MCS_MUR_PN, filename = "Figures/MCS.pNG")

# Here i get the years from 2013 to 2016
MCS <- MCS[11446:12541,]
MCS_OISST_PN <- ggplot(MCS, aes(x = t, y = temp)) +
geom_flame(aes(y2 = thresh), colour = "blue2", fill = "skyblue1") +
geom_text(aes(x = as.Date("2011-02-01"), y = 20,
              label = "PN_MCS")) +
  xlab("Date") + ylab(expression(paste("Temperature [", degree, "C]"))) +
  ggtitle("OISST")
ggsave(plot = MCS_OISST_PN, filename = "Figures/event_detect/PN/MCS_OISST_PN.pNG")



ts <- ts2clm(G1SSTsub_SST, climatologyPeriod = c("2002-06-01", "2015-12-30"), pctile = 10)
res <- detect_event(ts, coldSpells = TRUE)
MCS <- res$event

loli_G1SST_PN <- ggplot(MCS, aes(x = date_peak, y = duration)) +
  geom_lolli(aes(colour = intensity_cumulative)) +
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") +
  xlab("Date") + ylab("Event duration [days]")

#ggsave(plot = loli_G1SST_PN, filename = "Figures/event_detect/PN/loli_G1SST_PN.pNG")
```


```{r}
combined_flame <- ggarrange(MCS_SACTN_PN,MCS_OISST_PN,MCS_G1SST_PN, MCS_MUR_PN )
ggsave(plot = combined_flame, filename = "Figures/event_detect/PN/combined_flame.pNG")
```












